### 2022-01-26 cv/cv.tex
Add CV template

\documentclass[10pt, a4paper, oneside]{article}

\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=black, pdfauthor=Evgenii Iuliugin,
  pdftitle=Evgenii Iuliugin's CV}

\usepackage{indentfirst}
\usepackage[top=1.0cm, bottom=1.0cm, left=1.0cm, right=1.0cm,
  footskip=1.0cm]{geometry}

\usepackage{titlesec}
\titleformat
    {\section} % command
    [display] % shape
    {\normalfont\bfseries} % format
    {} % label
    {} % separator
    {\vspace{-2ex}} % before-code
    [
      \vspace{-2ex}
      \rule{\textwidth}{0.3pt}
    ] % after-code

\setlength\parindent{0pt}

\pagestyle{empty}

\begin{document}

\Huge
Evgenii Iuliugin

\LARGE
\textit{Software Engineer}

\begin{flushright}
  \large
  \href{mailto:yulyugin@gmail.com}{yulyugin@gmail.com}

  % phone number
\end{flushright}

\section*{SUMMARY}

\section*{EXPERIENCE}

\section*{EDUCATION}

\section*{PUBLICATIONS}

\section*{OTHER}

\end{document}