### 2022-02-08 cv/cv.tex
Fill in CV parts

\documentclass[12pt, a4paper, oneside]{article}

\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=black, pdfauthor=Evgenii Iuliugin,
  pdftitle=Evgenii Iuliugin's CV}

\usepackage{indentfirst}
\usepackage[top=1.0cm, bottom=1.0cm, left=1.0cm, right=1.0cm,
  footskip=1.0cm]{geometry}

\usepackage{titlesec}
\titleformat
    {\section} % command
    [display] % shape
    {\normalfont\bfseries} % format
    {} % label
    {} % separator
    {\vspace{-2ex}} % before-code
    [
      \vspace{-2ex}
      \rule{\textwidth}{0.3pt}
    ] % after-code

\setlength\parindent{0pt}

\pagestyle{empty}

\usepackage{tabularx}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}

\usepackage{enumitem}
\setlist[itemize]{noitemsep, topsep=-1em, leftmargin=2.5\labelsep}

\begin{document}

\Huge
\textbf{Evgenii Iuliugin}

\Large
\textit{Software Engineer}

\begin{flushright}
  \large
  \href{mailto:yulyugin@gmail.com}{yulyugin@gmail.com}

  % phone number
\end{flushright}

\normalsize

\section*{SUMMARY}

Software engineer with deep knowledge of hardware modeling and virtualization,
operating system internals, hardware architectures, and software design
principles. Result-oriented and well-qualified to develop complicated software
products from gathering technical requirements to full system integration
and debug.

Programming languages: C/C++ (10 years of experience); x86 assembly; Python.

Areas of interest: virtualization, operating systems, compiler technologies.

\section*{EXPERIENCE}

\begin{table}[htpb]
  \begin{tabularx}{\textwidth}{lY}
    2012 --- 2016 & \textbf{Intel Corporation} \hfill Moscow, Russia \\
    & \textbf{\textit{Software Engineer}} \\
    2016 --- now  & \textbf{Intel Corporation} \hfill Stockholm, Sweden \\
    & \textbf{\textit{Senior Software Engineer}} \\
    & \textit{Designed, implemented and optimized processor models. Integrated
    the models with virtual platforms used for pre-silicon validation.} \\
    & \begin{itemize}
      \item Led development of VT-x based acceleration in Intel Simics.
      \item Designed and implemented VT-x based acceleration in Intel Simics
        for macOS.
      \item Designed and implemeneted automatic generation of x86 instruction
        descriptions for Intel Simics using Intel XED instruction descriptions.
      \item Analyzed and improved performance of various Intel Simics processor
        and device models.
      \item Debugged and enabled boot scenario for different operating systems
        and hypervisors on Intel Simics.
      \item Collected and analyzed customer requirements, designed model
        components based on the requirements.
      \item Consulted team members on operating system internals.
      \item Regularly shared knowledge via technical presentations to the team.
      \item Regularly reviewed designs and code from peers.
      \item Mentored junior contributors.
      \end{itemize} \\
  \end{tabularx}
\end{table}

\section*{EDUCATION}

\begin{table}[htpb]
  \begin{tabularx}{\textwidth}{lY}
    2009 --- 2015 & \textbf{Moscow Institute of Physics and Technology} \hfill Moscow, Russia \\
    & \textit{MS in Applied Mathematics and Physics} \\
    2015 --- 2019 & \textbf{Moscow Institute of Physics and Technology} \hfill Moscow, Russia \\
    & \textit{PhD in Computer Science} \\
  \end{tabularx}
\end{table}

\section*{LANGUAGES}

\begin{itemize}
\item Russian --- fluent;
\item English --- intermediate;
\item Swedish --- basic.
\end{itemize}

\section*{SKILLS}

\begin{itemize}
\item Expert knowledge of virtual machine design and implementation,
  including binary translation and direct execution.
\item Knowledge of operating system design and implementation.
\item Deep knowledge of IA-32 and AMD64 processor architectures.
\item Cross-platform development for Windows, Linux and macOS with both native
  APIs and platform-agnostic libraries.
\item Advanced experience with driver development, including Windows Driver
  Model (WDM).
\item Collaborative development with Subversion, GIT, Jenkins, TeamCity, etc.
\item Proficiency with configuration management and build systems based on MAKE.
\end{itemize}

\section*{PUBLICATIONS}

\end{document}