### 2023-01-03 _posts/papers-2023.tex
Add short description for ericsson VP paper

\input{papers.tex}

\begin{document}

\section*{Simulators}

\begin{itemize}
    \item Ericsson's experience with virtual platform for pre-silicon software development. The virtual platforms are based on SystemC/TLM. This paper may be used as a showcase of simulation importance.

    \cite{Dahl:Ericsson-VP:2016}.
\end{itemize}

\end{document}