### 2023-01-14 paper-reviews/papers-2022.tex
Move paper reviews to a separate folder

\input{papers.tex}

\begin{document}

\section*{Machine Learning}
\begin{itemize}
    \item A new reconfigurable hardware architecture designed specifically to facilitate machine learning requirements:

    \cite{Prabhakar:Plasticine:2017}.

    \item An optimization to increase parallelism of machine learning tasks:

    \cite{Yuanzhong:Cross-Replica-Sharding:2020}.
\end{itemize}

\section*{Processors}
\begin{itemize}
    \item Western Digital builds its own RISC-V processor:

    \cite{Wheeler:WD-RISC-V:2019}.
\end{itemize}

\section*{Simulators}
\begin{itemize}
    \item Basic hardware performance modeling approach with separation of functional and timing simulation using ports:

    \cite{Emer:Asim:2002}.

    \item Micro architectural processor simulation framework:

    \cite{Hughes:Rsim:2002}.

    \item Infrastructure and framework for cycle-accurate and functional simulation:

    \cite{Austin:SimpleScalar:2002}.

    \item KVM-based direct execution for ARM:

    \cite{Junger:ARM-on-ARM:2020}.

    \item Overview of Simics simulator:

    \cite{Magnusson:Simics:2002}.

    \item Basic algorithm for multiprocessor target on multiprocessor host simulation:

    \cite{Kenney:MP-on-MP:2015}.

    \item Single instruction JIT compilation approach:

    \cite{Nohl:JIT-Simulation:2002}.

    \item Memory and thread usage analysis for software running inside a virtual platform:

    \cite{Dahl:VP-Sanitizer:2018}.
\end{itemize}

\section*{Compilers}
\begin{itemize}
    \item An amazing overview of basic but demonstrative optimizations performed by a modern C/C++ compiler:

    \cite{Godbolt:Optimizations:2020}.
\end{itemize}

\section*{Embedded Systems}
\begin{itemize}
    \item PhD thesis by Jakob Engblom focused on studying of worst case execution time analysis for embedded systems:

    \cite{Engblom:WCET:2002}.
\end{itemize}

\section*{Fun}
\begin{itemize}
    \item Really fun connection between texts of several popular songs and computer basics:

    \cite{Colwell:Rock-N-Roll:2002}.
\end{itemize}

\section*{Performance}
\begin{itemize}
    \item A way to determine maximum performance of a particular system together with few ideas on how to decide what optimizations make sense to do and when:

    \cite{Williams:Roofline:2009}.
\end{itemize}

\section*{Software Philosophy}
\begin{itemize}
    \item Discussions on what a perfect design process show look like:

    \cite{Parnas:Fake-It:1986}.

    \item Philosophical discussions about causes of errors in software:

    \cite{Horner:Error-Rates:2019}.
\end{itemize}

\section*{Virtual Machines}
\begin{itemize}
    \item A set of techniques to detect a virtualized Android environment using basic phone identification information, sensor data and timing of processor instruction execution:

    \cite{Petsas:Android-VM-Detection:2014}.
\end{itemize}

\end{document}