### 2025-02-20 paper-reviews/papers-2025.tex
Add titans paper

\input{papers.tex}

\begin{document}

A total of \total{citenum} papers read in the year.

\section*{Machine Learning}
\begin{itemize}
    \item \cite{Behrouz:Titans:2024}:

    The authors present Titans --- a new machine learning model architecture that incorporates long-term memory with forgetting in order to be able to handle very large context windows. A forgetting mechanism is added to the model to avoid memory overflows. Long-term memory is added as a context to the current input sequence.
\end{itemize}

\end{document}